---
title: "BK-CUSUM vs CGR-CUSUM"
author: "Daniel Gomon, Hein Putter, Rob Nelissen, Stéphanie van der Pas"
date: "30/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Clarification for the reader: initially we called the BK-CUSUM the CTCUSUM and the CGR-CUSUM was called the CTMAXGLR, the CGI-CUSUM was called the CTGLR.

#Load ALLFUNCTIONS.Rmd, contains all functions used in this script.

```{r, echo = FALSE}
rmarkdown::render("ALLFUNCTIONS.Rmd")
```

\section{Comparing the BK-CUSUM chart with the CGR/CGI-CUSUM chart}
In this file we will compare the BK-CUSUM chart with the CGR/CGI-CUSUM chart using $3000$ institutions with arrival rate $\psi = \frac{2500}{1095}$ with failure time distributed exponentially with rate $\lambda = 0.002$. We will consider the alternative hypothesis $\mu = \ln(1.4)$, where we construct the CUSUM with values $\theta \in \ln(\{1.1, 1.2, 1.3, 1.4, 1.5, 1.6\})$.

\subsection{Creating the 3000 in-control hospitals and out-of-control hospitals for each value of theta}

Under null-hypothesis $\mu = \ln(1) = 0$.
```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H0 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H0, file = "HOSPITALS3KEXP002H0.Rdata")
```

Under Alternative hypothesis $\mu = \ln(1.2)$
```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H12 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1.2))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H12, file = "HOSPITALS3KEXP002H12.Rdata")
```


Under Alternative hypothesis $\mu = \ln(1.4)$
```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H14 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1.4))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H14, file = "HOSPITALS3KEXP002H14.Rdata")
```


```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H16 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1.6))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H16, file = "HOSPITALS3KEXP002H16.Rdata")
```


```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H18 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1.8))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H18, file = "HOSPITALS3KEXP002H18.Rdata")
```



```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H20 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(2))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H20, file = "HOSPITALS3KEXP002H20.Rdata")
```

```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H22 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(2.2))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H22, file = "HOSPITALS3KEXP002H22.Rdata")
```

```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


HOSPITALS3KEXP002H24 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(2.4))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H24, file = "HOSPITALS3KEXP002H24.Rdata")

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
HOSPITALS3KEXP002H26 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(2.6))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H26, file = "HOSPITALS3KEXP002H26.Rdata")

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
HOSPITALS3KEXP002H28 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(2.8))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H28, file = "HOSPITALS3KEXP002H28.Rdata")

cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
HOSPITALS3KEXP002H30 <- foreach(j=1:3000) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  test2 <- gendatapoiss(psi = 2500/1095, t = 2000, coeffgen = coefficients, 
                         invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(3))
  test2
}
stopCluster(cl)
save(HOSPITALS3KEXP002H30, file = "HOSPITALS3KEXP002H30.Rdata")
```


  
\section{Determining control limits}
We would like to determine values for the control limits $h$ for the charts so that we can simulate detection speeds of both CGR/CGI- and BK-CUSUM charts.

\subsection{BK-CUSUM}
The "CTCUSUMcont" function keeps generating new data until at some point the control limit is surpassed. We therefore choose a control limit higher than the "true" control limit and determine the run length for $3000$ in-control hospitals using this control limit.
For the BK-CUSUM chart with $\theta = \ln(1.4)$ and $h = 8$ under the null hypothesis.
We take $h = 8$ as we know from running a few sample calculations that the final control limit will be smaller than $8$


```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


CONTCUSVOORH0THETA14h8 <- foreach(j=1:3000) %dopar% {
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  set.seed(j)
  a <- CTCUSUMcont(psi= 2500/1095, t = 1000, coeffsgen = c(0), invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1), gendata = gendatapoiss, coeffchart = c(0), cbaseh = function(t) chazexponential(t, lambda = 0.002), theta = log(1.4), h = 8)
  a
}
stopCluster(cl)
save(CONTCUSVOORH0THETA14h8, file = "E:/Scriptie/CONTCUSVOORH0THETA14h8.Rdata")
```

Same for $\theta = \ln(1.8)$, but now with control limit $h = 9$.


```{r, include = FALSE}
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


CONTCUSVOORH0THETA18h9 <- foreach(j=1:2000) %dopar% {
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  set.seed(j)
  a <- CTCUSUMcont(psi= 2500/1095, t = 1000, coeffsgen = c(0), invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1), gendata = gendatapoiss, coeffchart = c(0), cbaseh = function(t) chazexponential(t, lambda = 0.002), theta = log(1.8), h = 9)
  a
}
stopCluster(cl)
save(CONTCUSVOORH0THETA18h9, file = "E:/Scriptie/CONTCUSVOORH0THETA18h9.Rdata")
```

\subsection{CGI/CGR-CUSUM}

For the CGR-CUSUM chart with $h = 8$ under the null hypotesis.
"CTMAXGLRcont" is the analogue of "CTCUSUMcont".

```{r, include = FALSE}
#Not run, takes too long, split into different files
libary(foreach)
library(doParallel)
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)


CONTMAXGLRVOORH0h8 <- foreach(j=1:3000) %dopar% {
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  a <- CTMAXGLRcont(psi= 2500/1095, t = 1000, coeffsgen = c(0), invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1), gendata = gendatapoiss, coeffchart = c(0), cbaseh = function(t) chazexponential(t, lambda = 0.002),  h = 8)
  filenam <- paste0("D:/Scriptie/", "CTMAXGLR", as.character(j), ".Rdata" )
  #CGR-CUSUM runtime is quite long, so we save intermediary results.
  save(a, file = filenam)
  rm(nam)
  t <- 1
  return(t)
}
stopCluster(cl)
```







\section{Apply BK-CUSUM to data sets with alternative hypothesis failure rates.}

```{r}
hypoths <- c("12", "14", "16", "18", "20", "22", "24", "26", "28", "30")
for(i in 1:length(hypoths)){
  print(i)
  cores=detectCores()
  cl <- makeCluster(cores[1]-1) #not to overload your computer
  registerDoParallel(cl)
  load(paste0("HOSPITALS3KEXP002H", hypoths[i], ".Rdata"))
  hypoths <- c("12", "14", "16", "18", "20", "22", "24", "26", "28", "30")
  assign("test1", eval(parse(text = paste0("HOSPITALS3KEXP002H", hypoths[i]))))
  rm(list = eval(paste0("HOSPITALS3KEXP002H", hypoths[i])))
  Chartseval <- foreach(j=1:length(test1)) %dopar% {
    print("asd")
    coefficients <- c(1)
    chazexponential <- function(t, lambda){
      return(lambda * t)
    }
    invchazexponential <- function(t, lambda){
      return(t/lambda)
    }
    test <- test1[[j]]
    #We choose to perform simulations with about 2500 primary procedures per 1095 time points
    #The baseline hazard rate is assumed to be exponential with rate 0.002
    outpCUS <- lapply(log(c(1.4, 1.8)), function(x) CTCUSUMsim(test, coefficients = c(1), cbaseh = function(t) chazexponential(t, lambda = 0.002), theta = x, n = 2000, h = 9))
    outpCUS
  }
  assign(paste0("EvalChartsCUSUMH", hypoths[i]), Chartseval)
  save(list = eval(paste0("EvalChartsCUSUMH", hypoths[i])), file = paste0("EvalChartsCUSUMH", hypoths[i], ".Rdata"))
  stopCluster(cl)
  rm(Chartseval)
}
```

\section{Apply CGI-CUSUM to data sets with alternative hypothesis failure rates.}

```{r}
load("HOSPITALS3KEXP002H12.Rdata")
load("HOSPITALS3KEXP002H14.Rdata")
load("HOSPITALS3KEXP002H16.Rdata")
load("HOSPITALS3KEXP002H18.Rdata")
load("HOSPITALS3KEXP002H20.Rdata")
load("HOSPITALS3KEXP002H22.Rdata")
load("HOSPITALS3KEXP002H24.Rdata")
load("HOSPITALS3KEXP002H26.Rdata")
load("HOSPITALS3KEXP002H28.Rdata")
load("HOSPITALS3KEXP002H30.Rdata")
cores=detectCores()
cl <- makeCluster(cores[1]-1) #not to overload your computer
registerDoParallel(cl)
EvalChartsGLR <- foreach(j=1:length(HOSPITALS3KEXP002H14)) %dopar% {
  coefficients <- c(1)
  chazexponential <- function(t, lambda){
    return(lambda * t)
  }
  invchazexponential <- function(t, lambda){
    return(t/lambda)
  }
  test <- list(HOSPITALS3KEXP002H12[[j]],HOSPITALS3KEXP002H14[[j]],HOSPITALS3KEXP002H16[[j]],HOSPITALS3KEXP002H18[[j]],HOSPITALS3KEXP002H20[[j]],HOSPITALS3KEXP002H22[[j]],HOSPITALS3KEXP002H24[[j]],HOSPITALS3KEXP002H26[[j]],HOSPITALS3KEXP002H28[[j]],HOSPITALS3KEXP002H30[[j]])
  #We choose to perform simulations with about 2500 primary procedures per 1095 time points
  #The baseline hazard rate is assumed to be exponential with rate 0.002
  outpGLR <- lapply(test, function(x) CTGLRsim(x, coefficients = c(1), cbaseh = function(t) chazexponential(t, lambda = 0.002), n = 2000, h = 9))
  outpGLR
}
stopCluster(cl)
save(EvalChartsGLR, file = "EvalChartsGLR.Rdata")
```










\section{Looking at the ARL as a function of h under H0}

\subsection{BK-CUSUM}


```{r}
load("CONTCUSVOORH0THETA14h8.Rdata")
load("CONTCUSVOORH0THETA18h9.Rdata")
```

```{r}
hseq <- seq(5, 8, by = 0.01)
ARLh14 <- numeric(length(hseq))
for(i in 1:length(hseq)){
  runtimcusum <- sapply(CONTCUSVOORH0THETA14h8, function(x) RLfromh(x, h = hseq[i]))
  ARLh14[i] <- mean(runtimcusum)
}
hseq2 <- seq(5, 9, by = 0.01)
ARLh18 <- numeric(length(hseq))
for(i in 1:length(hseq2)){
  runtimcusum <- sapply(CONTCUSVOORH0THETA18h9, function(x) RLfromh(x, h = hseq2[i]))
  ARLh18[i] <- mean(runtimcusum)
}
```

```{r}
plot(hseq, ARLh14)
plot(hseq2, ARLh18)
```



\subsection{CGR-CUSUM}

```{r, include = FALSE}
CONTMAXGLRVOORH0h8 <- list()
for(i in 1:3000){
  load(file = paste0("E:/Scriptie/OLD/CTMAXGLRnieuw",as.character(i),".Rdata"))
  CONTMAXGLRVOORH0h8[[i]] <- a
  rm(a)
}

save(CONTMAXGLRVOORH0h8, file = "CONTMAXGLRVOORH0h8.Rdata")
```

```{r}
load("CONTMAXGLRVOORH0h8.Rdata")
```

```{r}
hseq <- seq(5, 8, by = 0.01)
ARLhMAXGLR <- numeric(length(hseq))
for(i in 1:length(hseq)){
  runtimMAXGLR <- sapply(CONTMAXGLRVOORH0h8, function(x) RLfromh(x, h = hseq[i]))
  ARLhMAXGLR[i] <- mean(runtimMAXGLR)
}
plot(hseq, ARLhMAXGLR)
plot(hseq, log(ARLhMAXGLR))
```

\subsection{Determining the values of h for ARL0 to be 15 years}
We would like to determine values of h for the BK-CUSUM and CGR-CUSUM so that ARL0 is 15 years. We do this as follows:
```{r}
year15 <- 365*15
hcus14 <- hseq[which.max(ARLh14 >= year15)] #6.82
hcus18 <- hseq2[which.max(ARLh18 >= year15)] #8.35
hMAXGLR <- hseq[which.max(ARLhMAXGLR >= year15)] #7.73
```




\section{Table for comparing ARL}

\subsection{BK-CUSUM}
```{r}
CUSrows <- data.frame(NULL,theta = double(), emu = double(), RL = double(), h = double())
hypoths <- c("12", "14", "16", "18", "20", "22", "24", "26", "28", "30")
hcus14 <- 6.82
hcus18 <- 8.35
hcus <- c(hcus14, hcus18)
for(k in 1:length(hypoths)){
  load(paste0("EvalChartsCUSUMH", hypoths[k], ".Rdata"))
  assign("test", eval(parse(text = paste0("EvalChartsCUSUMH", hypoths[k]))))
  rm(list = eval(paste0("EvalChartsCUSUMH", hypoths[k])))
  for(i in 1:length(test)){
    for(j in 1:length(test[[i]])){
      chart <- test[[i]][[j]]
      outp <- c("theta"= exp(chart$theta), "emu" = as.integer(hypoths[k])/10, "RL" = RLfromh(chart, h = hcus[j]), "h" = hcus[j])
      CUSrows <- rbind(CUSrows, outp)
    }
  }
  rm(test)
}
colnames(CUSrows) = c("theta", "emu", "RL", "h")
save(CUSrows, file = "CUSrows.Rdata")
```



\subsection{CGI-CUSUM}

```{r}
load("EvalChartsGLR.Rdata")
hmaxGLR <- 7.73
GLRrows <- data.frame(NULL, emu = double(), RL = double(), h = double())
emu <- seq(1.2, 3, by = 0.2)
for(i in 1:length(EvalChartsGLR)){
  for(j in 1:length(EvalChartsGLR[[i]])){
    chart <- EvalChartsGLR[[i]][[j]]
    outp <- c( "emu" = emu[j], "RL" = RLfromh(chart, h = hmaxGLR), "h" = hmaxGLR)
    GLRrows <- rbind(GLRrows,outp)
  }
}
colnames(GLRrows) = c( "emu", "RL", "h")
save(GLRrows, file = "GLRrows.Rdata")

```

\subsection{Under null hypothesis}

```{r}
hcus14 <- 6.82
hcus16 <- 7.67
hcus18 <- 8.35
hcus = c(hcus14, hcus16, hcus18)
CUSnullrows <- data.frame(NULL,theta = double(), emu = double(), RL = double(), h = double())
load("CONTCUSVOORH0THETA14h8.Rdata")
load("CONTCUSVOORH0THETA16h8.Rdata")
load("CONTCUSVOORH0THETA18h9.Rdata")
for(i in 1:length(CONTCUSVOORH0THETA14h8)){
  chart <- CONTCUSVOORH0THETA14h8[[i]]
  outp <- c("theta"= exp(chart$theta), "emu" = 1, "RL" = RLfromh(chart, h = hcus[1]), "h" = hcus[1])
  CUSnullrows <- rbind(CUSnullrows,outp)
}
for(i in 1:length(CONTCUSVOORH0THETA16h8)){
  chart <- CONTCUSVOORH0THETA16h8[[i]]
  outp <- c("theta"= exp(chart$theta), "emu" = 1, "RL" = RLfromh(chart, h = hcus[2]), "h" = hcus[2])
  CUSnullrows <- rbind(CUSnullrows,outp)
}
for(i in 1:length(CONTCUSVOORH0THETA18h9)){
  chart <- CONTCUSVOORH0THETA18h9[[i]]
  outp <- c("theta"= exp(chart$theta), "emu" = 1, "RL" = RLfromh(chart, h = hcus[3]), "h" = hcus[3])
  CUSnullrows <- rbind(CUSnullrows,outp)
}
colnames(CUSnullrows) = c( "theta", "emu", "RL", "h")
```

```{r}
hmaxGLR <- 7.73
GLRnullrows <- data.frame(NULL, emu = double(), RL = double(), h = double())
for(i in 1:length(CONTMAXGLRVOORH0h8)){
  chart <- CONTMAXGLRVOORH0h8[[i]]
  outp <- c( "emu" = 1, "RL" = RLfromh(chart, h = hmaxGLR), "h" = hmaxGLR)
  GLRnullrows <- rbind(GLRnullrows,outp)
}
colnames(GLRnullrows) = c( "emu", "RL", "h")
save(CUSnullrows, GLRnullrows, file = "nullrows.Rdata")
```


\section{Theoretical ARL}

Suppose we want to determine the ARL of CTCUSUM when $\mu = \ln(1.4)$ and $\theta \in \{\ln(1.2), \ln(1.4), \ln(1.6)}$ and the ARL of CTGLR when $\mu = \ln(1.4)$. We use exponentialy distributed failure times with rate 0.002 and $\psi = 2500/1095$.

\subsection{CTCUSUM}

```{r}
library(pracma)
omegaCTGLR <- function(omega, theta, mu){
 omega * (exp(theta) - 1) + exp(mu)*(exp(-omega*theta) - 1) 
}
diffomegaCTGLR <- function(omega, theta, mu){
  exp(theta) - 1 - theta*exp(mu)*exp(-omega*theta)
}
omega <- newtonRaphson(fun = function(x) omegaCTGLR(x, theta = theta, mu = mu), x0=-10, dfun = function(x) diffomegaCTGLR(x, theta = theta, mu = mu))$root

plotseq <- seq(-3, 10, by = 0.01)
plot(plotseq, omegaCTGLR(plotseq, theta = log(2), mu = log(1.22)))
```

```{r}
TheorARLCTCUSUM <- function(h, theta, mu, psi, cdfobs, C){
  #C is untill which time we consider the CUSUM
  Fi <- cdfobs(t = C, mu = mu)
  gammac <- exp(-mu)*psi * Fi
  nu = (theta * exp(mu) - exp(theta) + 1) * gammac
  if (nu == 0){
    return((h^2* exp(-mu))/(theta^2 * gammac))
  } else{
      library(pracma)
      omegaCTGLR <- function(omega, theta, mu){
        omega * (exp(theta) - 1) + exp(mu)*(exp(-omega*theta) - 1) 
      }
      diffomegaCTGLR <- function(omega, theta, mu){
        exp(theta) - 1 - theta*exp(mu)*exp(-omega*theta)
      }
      omega <- newtonRaphson(fun = function(x) omegaCTGLR(x, theta = theta, mu = mu), x0=100, dfun = function(x) diffomegaCTGLR(x, theta = theta, mu = mu))$root
      if(abs(omega) < 1e-4){
        omega <- newtonRaphson(fun = function(x) omegaCTGLR(x, theta = theta, mu = mu), x0=-100, dfun = function(x) diffomegaCTGLR(x, theta = theta, mu = mu))$root
      }
      finval <- h/nu - ((exp(theta - mu) - exp(-mu))/nu)*((1-exp(-omega*h))/(1-exp(-omega*theta)))
      return(c(finval, omega))
    }
}
```

We compare to the results written in Biswas \& Kalbfleisch

```{r}
a <- TheorARLCTCUSUM(h = 4.35, theta = log(2), mu = log(1.5), psi = 100, cdfobs = function(t, mu) cdfexp(t, mu,  lambda = -log(0.9)), C = 3)
b <- mean(replicate(50,CTCUSUMcont(psi= 100/365, t = 1000, coeffsgen = c(0), invchaz = function(t) invchazexponential(t, lambda = -log(0.9)/365), mugen = log(1.5), gendata = gendatapoiss, coeffchart = c(0), cbaseh = function(t) chazexponential(t, lambda = -log(0.9)/365), theta = log(2), h = 4.35)$runtime))/365
```

\subsection{CTGLR}

```{r}
#Use the fundamental theorem of calculus
TheorARLCTGLR <- function(h, mu, psi, intcdf, cdfobs){
  library(pracma)
  expecGLR <- function(t, mu, psi, intcdf, h){
    return((mu + exp(-mu) - 1)* psi * intcdf(t =t, mu = mu, lambda = 0.002) - h)
  }
  diffexpecGLR <- function(t, mu, psi, cdfobs){
    return((mu + exp(-mu) - 1)*psi * cdfobs(t, mu = mu, lambda = 0.002))
  }
  ARL <- tryCatch(newtonRaphson(fun = function(t) expecGLR(t, mu = mu, psi = psi, intcdf = intcdf, h=h), x0 = 10, dfun = function(t) diffexpecGLR(t, mu = mu, psi = psi, cdfobs = cdfobs), maxiter = 10000)$root, error = function(cond) return(Inf))
  return(ARL)
}
```

#Works well
```{r}
TheorARLCTGLR(h = 10,  mu = log(1.4), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp)
mean(replicate(30,CTGLRcont(psi= 2500/1095, t = 1000, coeffsgen = c(0), invchaz = function(t) invchazexponential(t, lambda = 0.002), mugen = log(1.4), gendata = gendatapoiss, coeffchart = c(0), cbaseh = function(t) chazexponential(t, lambda = 0.002),  h = 10)$runtime))
```

\subsection{CTCUSUMnew}
Theor ARL using our ARL for the CTCUSUM.

```{r}
#Use the fundamental theorem of calculus
TheorARLCTCUSUMv2 <- function(h, mu, theta, psi, intcdf, cdfobs){
  library(pracma)
  expecGLR <- function(t, mu, theta, psi, intcdf, h){
    return((theta + exp(-mu) - (exp(theta)/exp(mu)))* psi * intcdf(t =t, mu = mu, lambda = 0.002) - h)
  }
  diffexpecGLR <- function(t, mu, theta, psi, cdfobs){
    return((theta + exp(-mu) - (exp(theta)/exp(mu)))*psi * cdfobs(t, mu = mu, lambda = 0.002))
  }
  ARL <- tryCatch(newtonRaphson(fun = function(t) expecGLR(t, mu = mu, theta = theta, psi = psi, intcdf = intcdf, h=h), x0 = 10, dfun = function(t) diffexpecGLR(t, mu = mu, theta = theta, psi = psi, cdfobs = cdfobs), maxiter = 1000)$root, warning = function(cond){return(Inf)})
  return(ARL)
}
```




\section{Creating the Final Tables}

In this section we summarise above simulation results in a table.
```{r}
load("CUSrows.Rdata")
load("GLRrows.Rdata")
hcus14 <- 6.82
hcus16 <- 7.67
hcus18 <- 8.35
hmaxGLR <- 7.73
```

\subsection{For the CTCUSUM with $\theta = \ln(1.4)$}
```{r}
FinalTableCUS14 <- data.frame(NULL, ARL = double(), SD = double(), MRL = double(), TheorKB = double(), Theor = double())
tempdat <- subset(CUSrows, theta == 1.4)
for(i in unique(tempdat$emu)){
  temptemp <- subset(tempdat, emu == i)
  ARL <- round(mean(temptemp$RL))
  SD <- round(sd(temptemp$RL))
  MRL <- round(median(temptemp$RL))
  TheorKB <- round(TheorARLCTCUSUM(h = hcus14, theta = log(1.4), mu = log(i), psi = 2500/1095, cdfobs = function(t, mu) cdfexp(t, mu,  lambda = 0.002), C = 365)[1])
  Theor <- round(TheorARLCTCUSUMv2(h = hcus14,  mu = log(i), theta = log(1.4), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp))
  outp <- c( ARL, SD, MRL, Theor, TheorKB)
  FinalTableCUS14 <- rbind(FinalTableCUS14,outp)
}
colnames(FinalTableCUS14) = c("ARL", "SD", "MRL",  "Theor", "TheorKB (C=1)")
rownames(FinalTableCUS14) = unique(tempdat$emu)
```




\subsection{For the CTCUSUM with $\theta = \ln(1.8)$}
```{r}
FinalTableCUS18 <- data.frame(NULL, ARL = double(), SD = double(), MRL = double(), TheorKB = double(), Theor = double())
tempdat <- subset(CUSrows, theta == 1.8)
for(i in unique(tempdat$emu)){
  temptemp <- subset(tempdat, emu == i)
  ARL <- round(mean(temptemp$RL[is.finite(temptemp$RL)]))
  SD <- round(sd(temptemp$RL[is.finite(temptemp$RL)]))
  MRL <- round(median(temptemp$RL))
  TheorKB <- round(TheorARLCTCUSUM(h = hcus18, theta = log(1.8), mu = log(i), psi = 2500/1095, cdfobs = function(t, mu) cdfexp(t, mu,  lambda = 0.002), C = 365)[1])
  Theor <- round(TheorARLCTCUSUMv2(h = hcus18,  mu = log(i), theta = log(1.8), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp))
  outp <- c( ARL, SD, MRL, Theor, TheorKB)
  FinalTableCUS18 <- rbind(FinalTableCUS18,outp)
}
colnames(FinalTableCUS18) = c("ARL", "SD", "MRL",  "Theor", "TheorKB (C=1)")
rownames(FinalTableCUS18) = unique(tempdat$emu)
```

\subsection{For the CTGLR}
```{r}
FinalTableGLR <- data.frame(NULL, ARL = double(), SD = double(), MRL = double(), Theor = double())
for(i in unique(GLRrows$emu)){
  temptemp <- subset(GLRrows, emu == i)
  ARL <- round(mean(temptemp$RL))
  SD <- round(sd(temptemp$RL))
  MRL <- round(median(temptemp$RL))
  Theor <- round(TheorARLCTGLR(h = hmaxGLR,  mu = log(i), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp))
  outp <- c( ARL, SD, MRL, Theor)
  FinalTableGLR <- rbind(FinalTableGLR,outp)
}
colnames(FinalTableGLR) = c("ARL", "SD", "MRL", "Theor")
rownames(FinalTableGLR) = unique(GLRrows$emu)
```

```{r}
asd <- cbind(FinalTableCUS14, FinalTableCUS18, FinalTableGLR)
```

\section{Under the null hypothesis}

```{r}
load("nullrows.Rdata")
Finalrow <- numeric(0)
tempdat <- CUSnullrows
for(i in unique(tempdat$theta)){
  temptemp <- subset(tempdat, theta == i)
  ARL <- round(mean(temptemp$RL[is.finite(temptemp$RL)]))
  SD <- round(sd(temptemp$RL[is.finite(temptemp$RL)]))
  MRL <- round(median(temptemp$RL))
  TheorKB <- round(TheorARLCTCUSUM(h = temptemp$h[1], theta = log(i), mu = log(1), psi = 2500/1095, cdfobs = function(t, mu) cdfexp(t, mu,  lambda = 0.002), C = 365)[1])
  Theor <- round(TheorARLCTCUSUMv2(h = temptemp$h[1],  mu = log(1), theta = log(i), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp))
  outp <- c( ARL, SD, MRL, Theor, TheorKB)
  Finalrow <- c(Finalrow,outp)
}
temptemp <- GLRnullrows
ARL <- round(mean(temptemp$RL))
SD <- round(sd(temptemp$RL))
MRL <- round(median(temptemp$RL))
Theor <- round(TheorARLCTGLR(h = hmaxGLR,  mu = log(1), psi = 2500/1095, intcdf = integcdfexp, cdfobs =cdfexp))
outp <- c( ARL, SD, MRL, Theor)
Finalrow <- c(Finalrow,outp)

```

```{r}
Finalrow
```

asd and finalrow compose Table 3, the main simulation result of this thesis.










